```{r}
library(tidyverse)
library(dplyr)
library(lubridate)
```

Questions: 1) How do campaigns decide what bank to use?
2) What are the most popular banks used to finance elections?
3) What makes smaller banks, such as Chain Bridge Bank and Middletown Valley Bank, so popular among campaigns?
4) Is there correlation between location and bank used?
5) Have there been any significant trends in banks used by election campaigns in the last five years?
6) Are any large banks not as popular when used for campaign finance?
7) Are certain banks more “Republican” and certain banks more “Democratic”?

Reading in csv files from FEC website:

```{r}
bank_fees_24 <- read_csv("2024_Bank_Fees.csv")
bank_fees_23 <- read_csv("2023_Bank_Fees.csv")
bank_fees_22 <- read_csv("2022_Bank_Fees.csv")
bank_fees_21 <- read_csv("2021_Bank_Fees.csv")
bank_fees_20 <- read_csv("2020_Bank_Fees.csv")
bank_fees_19 <- read_csv("2019_Bank_Fees.csv")

bank_fees_24$recipient_zip <- as.character(bank_fees_24$recipient_zip)
bank_fees_23$recipient_zip <- as.character(bank_fees_23$recipient_zip)
bank_fees_22$recipient_zip <- as.character(bank_fees_22$recipient_zip)
bank_fees_21$recipient_zip <- as.character(bank_fees_21$recipient_zip)
bank_fees_20$recipient_zip <- as.character(bank_fees_20$recipient_zip)
bank_fees_19$recipient_zip <- as.character(bank_fees_19$recipient_zip)


```
First, I read in 6 years worth of data from the FEC website, going back to 2019. The data I read in specifically searched up the transaction "bank fees" that were reported to the FEC. I thought this would be the most efficient way to find out which campaigns used which banks. Naturally, some transactions were marked as bank fees which had nothing to do with banking at all, so I went through the data and removed everything that wasn't a bank, including merchant services, credit cards, credit unions, as well as transactions like Venmo and Zelle. Finally, I removed rows from the dataframe that weren't relevant to my project.


```{r}
core_data <- bind_rows(bank_fees_24, bank_fees_23, bank_fees_22, bank_fees_21, bank_fees_20, bank_fees_19)

core_data <- core_data %>% select(-image_number)
core_data <- core_data %>% select(-line_number)
core_data <- core_data %>% select(-entity_type_desc)
core_data <- core_data %>% select(-unused_recipient_committee_id)
core_data <- core_data %>% select(-recipient_committee_id)
core_data <- core_data %>% select(-national_committee_nonfederal_account)
core_data <- core_data %>% select(-memo_code_full)
core_data <- core_data %>% select(-original_sub_id)
core_data <- core_data %>% select(-back_reference_transaction_id)
core_data <- core_data %>% select(-back_reference_schedule_id)

core_data <- core_data %>% select(-starts_with("candidate_"))

core_data <- core_data %>% mutate(disbursement_date=mdy_hm(disbursement_date))

core_data <- core_data %>% mutate(bank_clean=recipient_name)

#step 1-remove non banks
#step 2-standardize

core_data <- core_data %>% filter(!recipient_name %in% c("ACTBLUE", "ANEDOT", "YU, EUGENE CHIN", "ZAHRADKA, TAWNJA", "VINCENT, ZACHARY", "VEED PRO", "US ONCOLOGY", "UNITEMIZED, EXPENSES", "UBER", "TOWNSEND FOR CONGRESS", "TARGETED VICTORY LLC", "TARGETED VICTORY", "STRIPE INC", "STRATEGIC PARTNERS & MEDIA LLC", "SOUTHERN IMPACT GROUP LLC", "SMALL UN-ITEMIZED EXPENDITURES", "SHAUN BROWN FOR VIRGINIA", "RYAN KRAUSE FOR CONGRESS", "SIFMA", "WILKINS STRATEGIES GROUP, LLC", "WESTERN ALLIANCE", "WALSH, SEAN P", "SOUTH MAIN STREET PROPERTIES II LTD", "SIGNAPAY", "RIGHT WOMEN PAC", "RAMOS, JAVIER", "PLUG & PLAY TECHNOLOGIES, INC.", "OLIVER, CHASE", "NORPAC", "NEW BLUE INTERACTIVE LLC", "NEW BLUE INTERACTIVE, LLC", "NATIONAL PROPANE GAS ASSOCIATION", "NCPSSM", "MITCHELL, CAITLIN", "MESSER CONSTRUCTION CO PAC", "MARSH AND MCLENNAN AGENCY LLC", "MARILYN FOR U S SENATE", "LEONID GRINBERG LLC", "K&L GATES LLP", "JUST REFORM PAC", "JONES, HUGH", "JANICEK FOR SENATE", "IMAGE DIRECT", "I. LANIER AVANT", "HUBBARD, LOURIN", "HONIGMAN, WILLIAM", "HOADLEY HOUSE LLC", "HENRYALAN", "GLOBAL SERVICE FEE", "GLOBAL PAYMENTS", "FRIENDS OF MICHAEL HOOD", "FOUTS, MATT", "GILSTRAP INC", "GILLERMAN, JONATHAN", "EZELL, MILTON LEVELL", "ELIAS LAW GROUP", "ELITEWORK PROPERTIES LLC", "EDUCATOR'S CU", "DONORBOX", "DIRECT MAIL PROCESSORS", "DIRECT MAIL PROCESSORS", "DEMATTEIS & RICCIARDELLO PLLC", "DC LOCKSMITH", "DANI FOR ILLINOIS", "DALSIMER, ALISON", "CONYERS, JOHN III", "CHICK-FIL-A", "CERTIFIED CONSERVATIVES PAC", "CALL HUB - GAGLERS INC.", "BRITEY LLC", "BROOKS FOR SENATE", "BRAINTREE", "BP BUSINESS CARD", "BILLHIGHWAY", "BIDPAL", "AUTHORIZENET", "AT&T", "ARK MED SOC PAC", "ARK MED SOCIETY PAC", "AMS POLITICAL ACTION COMMITTEE", "ACT BLUE AND BANK FEES", "10SIX CONSULTING", "WINRIGHT SRATEGIES, LLC", "WEBSITE HOSTING SERVICE", "VERITAS PROGRESSIVES PAC", "VANTIV ECOMMERCE", "TRUJILLO, CRYSTAL J DR.", "TRI COUNTIES", "THRIPP, RICHARD", "THREE ARBOR INSURANCE INC.", "THE GIVING BLOCK", "STEPHENS FOR CONGRESS", "SPEEDPRO IMAGING", "SALEM MEDIA GROUP", "RURAL ARIZONA ACTION", "REVV.CO DBA WINRED", "REVV FUNDRAISING PLATFORM", "RENASANT", "RAISE THE MONEY, INC.", "RAISE THE MONEY", "P & A ADMINISTRATIVE SERVICES INC", "PEX CARD", "NMI", "NGP VAN, INC.", "MOONCLERK", "LUCY LAUSER FOR CONGRESS", "JON CURTIS FOR CONGRESS", "HUDSON BAY COMPANY OF ILLINOIS", "HINDUS OF GEORGIA PAC", "GOODMAN CAMPAIGNS LLC", "GOOGLE", "VERGEBEN SERVICES COMPANY", "ACTBLUE TECHNICAL SERVICES", "ELAVON", "BLUEPAY, INC.", "SHOPIFY INC", "WINRED", "WEPAY, INC.", "OREGON RIGHT TO LIFE", "ALLIANT MERCHANT SOLUTION", "ANEDOT, INC.", "SHOPIFY INC.", "CIBC", "NUSENDA CREDIT UNION", "HEARTLAND PAYMENT SYSTEMS", "NUMERO", "CITI CARDS", "PEX", "PAYA", "FIRST BANKCARD", "BANKCARD ASSOCIATES", "CARDMEMBER SERVICE", "ALLIANT LLC", "UBT", "VOTESANE PAC", "HANCOCK WHITNEY", "CORNERSTONE PAYMENT SYSTEMS", "CARL, JERRY LEE JR", "TEXAS MEDICAL ASSOC", "CARDMEMBER SERVICES", "AXIA BANKCARD MERCHANT", "BOK FINANCIAL", "SQUARE, INC", "LOCAL 94-94A-94B IUOE - UNION", "ALLIANT MERCHANT SERVICES", "DHARMA MERCHANT SERVICES", "PINNACLE FINANCIAL PARTNERS", "STRIPE, INC.", "CHASE CARD SERVICES", "PARAGON PAYMENT SOLUTIONS", "ONEAZ CREDIT UNION", "AMERISERV FINANCIAL", "AMERICAN EXPRESS BANK FEES", "ACT BLUE", "UMB", "MERCHANT CARD SERVICES", "HUDSON VALLEY CREDIT UNION", "DEMOCRACY ENGINE", "CHASE CARDMEMBER SERVICES", "BEST CARD", "BALL STATE FEDERAL CREDIT UNION", "AUTHNET GATEWAY", "AA SERVICE FEES", "VANCO", "INTEGRATED SOLUTIONS", "FEDWIRE", "CYBERSOURCE", "WORLD PAY", "PAYPAL HOLDINGS, INC.", "AMERICAN EXPRESS COMPANY", "INTUIT PAYMENT SOLUTIONS", "ELITE CARD PROCESSING LLC", "DEMOCRACY ENGINE, LLC", "CITI CARD", "C&H FINANCIAL SERVICES", "AUTHNET", "BARCLAY'S CREDIT CARD", "SQUARE, INC.", "HUNTINGTON", "HONOR CREDIT UNION", "CHOICE FINANCIAL", "TRUIST MERCHANT SERVICES", "OMEGA PROCESSING", "INTUIT", "MARKETAXESS HOLDINGS INC. PAC", "FIRST DATA", "VISA", "VERGEBEN SERVICES", "WINRED TECHNICAL SERVICES") 
                  )

core_data <- core_data %>% filter(!recipient_name %in% c("EDONATION.COM", "PAYPAL MERCHANT SOLUTIONS, NA", "SQUARE", "SQUARE INC", "VENMO", "PAYPAL", "AUTHORIZE NET", "ELEVON", "V24-013", "PIRYX, INC.", "DNC TRAVEL ESCROW", "ISTREAM", "ENGAGE POLITICS", "ELITE CARD PROCESSING", "WORLDPAY, INC", "TRANSAXT", "SQUARESPACE", "SQUARE FEES", "RALPH AND LOIS DREW TRUST", "SAFEGUARD", "PUBLIX", "PRO ISRAEL AMERICA PAC", "OREGON RIGHT TO LIFE - C4", "MAVERICK PAYMENTS (M - MERCHANT)", "INFINEX INVESTMENTS INC.", "FULLSTREAM", "GEMINI", "DUPACO", "DELUXE CORPORATION", "DELUXE", "DONORBUREAU", "COOPERATIVE FEDERAL", "COMCAST CABLE", "COMMITTEE TO ELECT NATE CAIN", "CIGAR-PAC", "ALLY WESTSHORE LLC", "AFFINIPAY", "TRANSFIRST HOLDINGS", "SOUTHWESTERN ILLINOIS LECET", "HSP", "HARLAND CLARKE CHECKS", "FIRSTDATA MERCHANT SERVICES", "DROPBOX", "CLOVER.COM", "COMMERCIAL CREDIT CARDS", "BANK,, WEBSITE FEES", "AMERICAN EXPRESS SKYE BLUE", "AMERICAN EXPRESS GOLD 2", "UNITED STATES POSTAL SERVICE", "STRIPE, INC", "TD BUSINESS CARD", "SIMMONS BANK CREDIT CARD #4514", "STRIKE WEALTH INVESTMENT", "PAYROLL DATA PROCESSING INC", "MASTERCARD SERVICE CENTER", "MERCHANT SERVICE FEES", "MASTERCARD", "FULLSTEAM", "COGENCY GLOBAL INC", "ELAVON INC.", "EAGLE CREDIT CARD", "DISCOVER IT CARD", "CAMPAIGN FINANCIAL SERVICES", "BENNET COLORADO VICTORY FUND", "BB&T BANK CARD", "CHASE BA VISA - SKYE", "CITYWIDE", "CITI PRESTIGE", "ARISTOTLE INTERNATIONAL, INC", "AMERICAN EXPRESS CARD SERVICES", "AMERICAN EXPRESS BLUE BUS+", "USPS", "9SEVEN CONSULTING", "ACTBLUE TECHINICAL SERVICES", "AADVANTAGE MASTERCARD", "STATION SQUARE LLC", "QUICKBOOKS MERCHANT SERVICES", "QUICKBOOKS MERCHANT SERVICES", "FRIENDS OF KIRBY BIRGANS", "INTUIT, INC.", "OPPENHEIMER & CO. INC.", "MSCPAC", "FEES, BANK", "ELITE CARD PRO", "FIRST DATA MERCHANT SERVICES, LLC", "CARDMEMBER SERVICES - CREDIT CARD", "BENESYS, INC.", "BRAND NEW SENATE", "CARD CONNECT", "CITI BUSINESS CARD", "BANK FEES", "APWU", "WIN RED TECHNICAL SERVICES LLC", "VERIDIAN", "SCHWAB-HELLER", "VANTIV ECOMMERCE FUNDS", "QR-CODE-GENERATOR.COM", "PARAGON SOLUTIONS", "PAYPAL FEES", "NON BB CARD NOT PRESENT", "HOLYK, ISAAC", "INTELLZ, INC.", "MERCHANT CARD PROCESSING", "EVERTEC, INC", "CFS COMPLIANCE", "CARD SERVICES", "STIKE WEALTH INVESTMENT", "SENTINEL FEDERAL CREDIT UNION", "SECURITY BANKCARD CENTER, INC.", "SEFCU", "PARAGON PAYMENT SOLUTIONS, INC.", "PARAGON PAYMENT SOLUTION", "INTUIT, INC", "INTUIT MERCHANT SERVICES", "EXPRESS, AMERICAN", "CREDIT UNION NATIONAL ASSOCIATION", "CHASE CREDIT CARD", "CITIBANK CITI CARD", "COMMODORE PERRY FCU", "CAMPAIGN FINANCIAL SERVICES (CFS)", "BALL STATE FEDERAL CU", "VSECU", "WESTERN UNION", "UTILITIES EMPLOYEES CREDIT UNION", "PATRIOT FEDERAL CREDIT UNION", "FIRST VIRGINIA COMMUNITY-MADERA", "FIRST VIRGINIA", "DEMOCRACY ENGINE LLC", "COMMERCE BANKSHARES INC.", "COBALT CREDIT UNION VISA (SAC FCU)", "BRITISH AIRWAYS CHASE", "COBALT CREDIT UNION", "WSECU", "UW CREDIT UNION", "TRANSFIRST LLC", "HUBAY LLC", "RED CURVE SOLUTIONS", "DUNHAM TRUST COMPANY", "REVV LLC", "CHASE CARD BRITISH AIRWAYS", "BRENIZER, TAMMY", "BEST CARD, LLC", "WIRE TRANSFER CHARGE", "SQUARESPACE INC.", "SIGNATURE FCU", "STRIPE PAYMENT PROCESSING", "SERVISFIRST", "PAYPAL, INC.", "MB FINANCIAL", "POLITICAL COMMUNICATIONS, INC", "NORTH ISLAND CREDIT UNION", "NATIONBUILDER", "PEAK BUSINESS SOLUTIONS", "CITICARDS", "BANK OF AMERICA CREDIT CARD", "BB&T/TRUIST CREDIT CARD", "TSYS", "HAZE FOR CONGRESS COMMITTEE", "HARLAND CLARKE", "GIVESMART - COMMUNITY BRANDS", "BILL.COM", "AXCAPITAL, LLC", "WELLS FARGO CREDIT CARD", "SYNOVUS", "PROPAY INC", "POLITICAL COMMUNICATIONS, INC.", "MECU CREDIT UNION", "HOME FEDERAL", "EVENTBRITE", "ENT CREDIT UNION", "CORPORATE ONE FCU, INC.", "CLASSIC CITY", "CARD SERVICE CENTER", "C SPIRE", "BB&T BANK CREDIT CARD", "BANK SERVICE FEES", "NGP FEES", "ROYAL CREDIT UNION", "NGP FEES", "CHASE CARD SERVICE", "BANKCARD USA MERCHANT FEES", "STRIPE", "MELIO", "NGP VAN", "AUTHORIZE.NET", "AMERICAN EXPRESS", "ANEDOT.COM", "ARISTOTLE INTERNATIONAL", "SYNOVUS/NBSC", "MERCHANT SERVICES") 
)

core_data <- core_data %>% filter(!recipient_name %in% c("SHOPIFY", "SUPERIOR CREDIT UNION, INC.", "MERCHANT BANK CARD", "FIRSTBANK MERCHANT SERVICES", "ASSOCIATED CREDIT SERVICES, INC.", "ALLOYA CORPORATE - RAFFLE ACCOUNT", "THE BILLY PREMPEH FOR CONGRESS COMMITTEE", "SIMMONS BANK CREDIT CARD #2904", "PNC MERCHANT SERVICES", "MICHIGAN CREDIT UNION LEAGUE - VENDOR", "ALLOYA CORPORATE CREDIT UNION - RAF", "AMALGAMATED BANK CREDIT CARD", "U S BANK MERCHANT FEES", "PNC MERCHANT", "COMMERCE BANK CREDIT CARDS", "COMMITTEE TO ELECT KIMBERLY BIZON", "HARLAND CLARKE CORP.", "USAEPAY", "SIGNATURE CHECKING", "AMERICAN EXPRESS MERCHANT SERVICES", "BANK OF AMERICAN MERCHANT SERVICES", "AMALGAMATED INVESTMENT", "BANK FEES, N/A", "MILLBURY FEDERAL CREDIT UNION", "FIRST BANK CHICAGO - CARDMEMBER SERVICE", "CHASE PAYMENTECH", "CHASE PAYMENT TECH", "PAYPAL INC.", "BARCLAYS", "AADVANTAGE AVIATOR RED MASTERCARD", "BANKCARD", "EVENTBRITE INC", "FIRST BANK CARD", "STREAMLINE PAYMENTS", "SUNTRUST MERCHANT SERVICES", "CITIBANK CBO SERVICES", "TD MERCHANT SERVICES", "FIRST BANK MERCHANT SERVICES", "MORGAN STANLEY SMITH BARNEY LLC", "SPOKANE TEACHERS CREDIT UNION", "SOUTH CAROLINA FEDERAL CREDIT UNION", "BANKCARD CENTER", "CARDCONNECT", "ALLIANT MERCHANT SOLUTIONS", "STRIPE INC.", "TOWN AND COUNTRY CREDIT UNION", "FIRST DATA MERCHANT SERVICES")
)

write_csv(core_data %>% distinct(recipient_name, bank_clean), "banks_for_open_refine.csv")

```

I found the most commonly used banks according to my dataset: 

What's interesting here is that the top two banks, as well as two others in the top 10, are not known major financial institutions.

According to my data, the number one bank used by political campaigns is Amalgamated Bank. With just 5 branches across D.C., New York, and San Francisco, Amalgamated Bank is clearly not a household bank when it comes to consumers. According to many in D.C., Amalgamated Bank is known as the "private bank" for the Democratic party. They hold close affiliation with major labor unions, progressive activists, and top democratic groups in D.C. In total, Amalgamated Bank holds about 7 billion dollars in deposits, which is hardly anything compared to the 443 billion held by PNC Bank. It's clear Amalgamated Bank serves a need, and it serves it well.

The number two bank used by political campaigns is Chain Bridge Bank. It serves a very similar purpose to Amalgamated Bank, except for the Republican party. With just a singular branch in McLean, Virginia, Chain Bridge Bank holds just under 2 billion dollars in deposits. They are known as the preferred bank of the Republican party. They have relationships with several top Republican committees associated with Trump, Romney, and McCain, as well as with large conservative super-PACs, lobbying firms, and the Republican National Committee.

The other two banks in the top 10 that aren't major U.S. banks are First Virginia Community Bank and Middletown Valley Bank. 

First Virginia Community Bank, or FVC Bank, is a regional bank with 44 branches across the east coast.  Middletown Valley Bank hasabout 10 branches in Maryland and Pennsylvania. Both of these banks are located fairly close to D.C. and seem to specialize in campaign financing, however neither bank really appears to be affiliated with any particular party. 

```{r}
core_data %>% 
  group_by(recipient_name) %>% 
  summarize(transactions = n()) %>% 
  arrange(desc(transactions))
```

If I had more time on this project, I would use the cleaned up dataset to dive deeper into the banks used in campaign financing. I would see what correlations there are between campaigns and banks in different states, for example if PNC is the popular bank among PA- based campaigns, and which states favor Bank of America or Chase. I'd also like to look into Middletown Valley and FVC bank more, and see what kinds of campaigns use those banks.

```{r}
clean_banks <- read_csv("BANK_PROJ.csv") 
  
new_banks <- core_data %>% left_join(clean_banks) 

group_by (recipient_name) %>% 
  summarize (transactions = n())

```



