---
title: "lab_02"
author: "derek willis"
date: "2024-08-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

For this lab we'll be working with a slice of Maryland voter registration data: records of people who have registration dates this year through Aug. 7. Let's load it.

```{r}
md_voters_2024 <- read_rds("data/lab_02.rds")
```

## Examine Data

Take a look at the data, getting familiar with what it contains. Most of the columns should be self-explanatory, but there are two date columns: STATE_REGISTRATION_DATE and COUNTY_REGISTRATION_DATE. The former represents the date the person registered as a Maryland voter; the latter is when they registered in their current county. They may be the same, but if a voter moves from one county to another they will differ. This is important because if the two dates are the same, that's a pretty good indication that this is a *new* voter, someone who has no history of registering in Maryland. You'll need to keep that definition in mind.

## Answer questions

**Q1** How many voters appear to be *new* registrants since July 21, when President Joe Biden announced he would leave the race, based on the information above? Create a dataframe to hold the only those voters. Which counties have seen the most? Take a look at county population figures here and compare the two in your answer: https://worldpopulationreview.com/us-counties/maryland.

**17,083 new voters since July 21**

```{r}
new_voters <- md_voters_2024 |>
  filter(STATE_REGISTRATION_DATE > "2024-07-21")

new_voters %>% 
  group_by(COUNTY) %>% 
  summarise(voters=n()) %>% 
  arrange(desc(voters))



```

**Q2** A better question is whether the new voters you found in Q1 are an outlier compared to other weeks this year. Using the original dataframe with 2024 registrants, use `mutate()` and lubridate to produce week numbers from the COUNTY_REGISTRATION_DATE (Lubridate has a function that turns a date into a week number. Use that to answer this question). Save that to the original dataframe. Then calculate how many voters are in each week. Describe the distribution of voters groups in a sentence - are the weeks from July 21-August 7 different from the rest, and if so, how?

**Yes, the weeks of 7/21 and 7/28 account for 2 of the top 4 most popular weeks of registration in 2024.**

```{r}
md_voters_2024 %>% 
  mutate(week = floor_date(STATE_REGISTRATION_DATE, "week")) %>% 
  group_by(week) %>% 
  summarise(voters=n()) %>%
  arrange(desc(voters))

```

**Q3** Using lubridate, show which month saw the largest number of registrations. Does the answer help shed more light on Q2?
**July shows the most registrations. This was right around the Trump assassination attempt and Biden dropping out. Yes, this sheds more light on Q2, July was a very active month for voter registration.** 

```{r}
md_voters_2024 %>% 
  mutate(month = floor_date(STATE_REGISTRATION_DATE, "month")) %>% 
  group_by(month) %>% 
  summarise(voters=n()) %>%
  arrange(desc(voters))
```

**Q4** Maryland has, officially, five regions: https://www.visitmaryland.org/article/maryland-regions. Modify the original dataframe again to add a `region` column based on the `COUNTY` value using case_when. Are you going to have to do this for every Maryland county plus Baltimore City? Yes, you are. Which regions and parties account for the most new voters?

**The Democratic Party accounts for the most new voters, with many of them coming from Central Maryland.** 

```{r}
md_voters_2024 <- md_voters_2024 %>% 
   mutate(region = case_when(
        COUNTY == "Allegany" ~ "Western Maryland",
        COUNTY == "Garrett" ~ "Western Maryland",
        COUNTY == "Washington" ~ "Western Maryland",
        COUNTY == "Frederick" ~ "Capital",
        COUNTY == "Montgomery" ~ "Capital",
        COUNTY == "Prince George's" ~ "Capital",
        COUNTY == "Anne Arundel" ~ "Central Maryland",
        COUNTY == "Baltimore City" ~ "Central Maryland",
        COUNTY == "Baltimore County" ~ "Central Maryland",
        COUNTY == "Carroll" ~ "Central Maryland",
        COUNTY == "Harford" ~ "Central Maryland",
        COUNTY == "Howard" ~ "Central Maryland",
        COUNTY == "Calvert" ~ "Southern Maryland",
        COUNTY == "Charles" ~ "Southern Maryland",
        COUNTY == "St. Mary's" ~ "Southern Maryland",
        COUNTY == "Kent" ~ "Eastern Shore",
        COUNTY == "Queen Anne's" ~ "Eastern Shore",
        COUNTY == "Talbot" ~ "Eastern Shore",
        COUNTY == "Caroline" ~ "Eastern Shore",
        COUNTY == "Dorchester" ~ "Eastern Shore",
        COUNTY == "Wicomico" ~ "Eastern Shore",
        COUNTY == "Somerset" ~ "Eastern Shore",
        COUNTY == "Worcester" ~ "Eastern Shore",
        COUNTY == "Cecil" ~ "Eastern Shore"
    ))
          
md_voters_2024  %>%  
  group_by(region, PARTY)  %>%  
  summarize(voters = n()) %>% 
  arrange(desc(voters))

```

**Q5** What's an additional question you could ask of this data?

**Another question you could ask this data might be "How many registered republicans are there in the state of Maryland?"** 
